19Q1

### 2019-05-09 19Q1.Rmd
update Rmd

```{r load_drug_rds, eval=TRUE}
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

# Testing
test_ccle <- ccle_ptmuts %>% count(Gene, Mutation_Status)
test_ctrp <- ctrp_ptmuts %>% count(Gene, Mutation_Status)
test_gdsc <- gdsc_ptmuts %>% count(Gene, Mutation_Status)
```

@@

Merge all CRISPR genes:
```{r merge_crispr_data, eval=TRUE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

@@

# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Gene) %>% summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1), N_Mutant = sum(Mutation_Status == 1))
## Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Gene")
crispr_ptmuts$Mutation_Status <- factor(crispr_ptmuts$Mutation_Status)
## Add Color column
crispr_data$Color <- ifelse(crispr_data$Mutation_Status == 0, "thistle", "darkorchid")
crispr_data$Color <- factor(crispr_data$Color)

@@

```{r ccl_venn, eval=TRUE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

@@

```{r drug_venn, eval=TRUE}
ccle_drugs <- unique(ccle_ptmuts$Drug)
ctrp_drugs <- unique(ctrp_ptmuts$Drug)
gdsc_drugs <- unique(gdsc_ptmuts$Drug)

@@

```{r histology_dists_data, eval=TRUE}
ccle_hists <- unique(select(ccle_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

@@

```{r, eval=TRUE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

@@

```{r wilcoxon_tests, eval=TRUE}
crispr_ptmuts <- unfactorize(crispr_ptmuts)
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

@@

```{r mut_heatmaps, eval=TRUE}
ht_ccle <- MakeMutHeatmap(signif_g2p_druggene = ccle_signif_g2p_druggene, legend_title = "P-value", row_names = TRUE, color = "darkgreen", col_title = "Drugs")
png(filename = "./plots/manuscript/ccle_mut_heatmap.png", width = 3, height = 10, units = "in", res = 500)
draw(ht_ccle, heatmap_legend_side = "top")
dev.off()

@@

```{r drug_spear_calc, warning=FALSE, eval=TRUE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

@@

```{r drug_spear_plots, eval=TRUE}
# Copy number
ccle_cor <- unfactorize(filter(g2p_cor, Dataset == "CCLE"))
ccle_cn_ht <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.25, 0, 0.25), legend_title = "CCLE Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/ccle_cn_heatmap.png", width = 3, height = 10, units = "in", res = 500)
draw(ccle_cn_ht, heatmap_legend_side = "top")
dev.off()

@@

ccle_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
ctrp_ge_ht2 <- MakeCorrHeatmap(druwg_cor_res = ctrp_cor, row_names = FALSE, col_title = "CTRP", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
gdsc_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = gdsc_cor, row_names = FALSE, col_title = "GDSC", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
png(filename = "./plots/manuscript/drug_ge_heatmap.png", width = 13, height = 10, units = "in", res = 500)
draw(ccle_ge_ht2 + ctrp_ge_ht2 + gdsc_ge_ht2, gap = unit(0.25, "in"), heatmap_legend_side = "top")
dev.off()
```

@@

Kolmogorov-Smirnov tests comparing DGLs that are in a G2P DGL and not.
```{r drug_KS_tests, eval=TRUE, warning=FALSE, fig.height=6, fig.width=12}
g2p_dgl_corr <- G2PCorrTable(ccle_ptmuts_df = ccle_ptmuts, ctrp_ptmuts_df = ctrp_ptmuts, gdsc_ptmuts_df = gdsc_ptmuts)
saveRDS(g2p_dgl_corr, "./data_munging/rds/g2p_dgl_corr.rds")
g2p_dgl_corr <- readRDS("./data_munging/rds/g2p_dgl_corr.rds")
g2p_mut_stats <- G2PCorrStatsTables(corr_df = g2p_dgl_corr, table_type = "mutation")

g2p_corr_4plot <- filter(g2p_corr_all, G2P_Drug_Gene == "Yes")
g2p_corr_stats <- G2PCorrStatsTables(corr_df = g2p_corr_4plot, table_type = "KS")
g2p_corr_4plot$p <- NULL
g2p_corr_4plot$Mut_Signif <- NULL
g2p_corr_4plot <- melt(g2p_corr_4plot, id.vars = colnames(g2p_corr_4plot)[1:6], measure.vars = colnames(g2p_corr_4plot)[7:8], value.name = "Correlation")
g2p_corr_4plot <- g2p_corr_4plot %>% separate(col = variable, into = c("Correlation_Method", "corr", "Metric"), sep = "_", remove = FALSE)
g2p_corr_4plot$Metric <- ifelse(g2p_corr_4plot$Metric == "GE", "Gene Expression", "Copy Number")

>>>>

```{r load_drug_rds, eval=TRUE}
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds") %>% drop_na(Drug)
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds") %>% drop_na(Drug)
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds") %>% drop_na(Drug)

# Testing
test_ccle <- ccle_ptmuts %>% count(Gene, Drug)
test_ctrp <- ctrp_ptmuts %>% count(Gene, Mutation_Status)
test_gdsc <- gdsc_ptmuts %>% count(Gene, Mutation_Status)
```

@@

Merge all CRISPR genes:
```{r merge_crispr_data, eval=FALSE}
crispr_all <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr_all) <- gsub(" .*", "", colnames(crispr_all))
colnames(crispr_all)[1] <- "Broad_ID"

@@

# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Gene) %>% summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1), N_Mutant = sum(Mutation_Status == 1))
## Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Gene")
crispr_data$Mutation_Status <- factor(crispr_data$Mutation_Status)
## Add Color column
crispr_data$Color <- ifelse(crispr_data$Mutation_Status == 0, "thistle", "darkorchid")
crispr_data$Color <- factor(crispr_data$Color)

@@

```{r ccl_venn, eval=FALSE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

@@

```{r drug_venn, eval=FALSE}
ccle_drugs <- unique(ccle_ptmuts$Drug)
ctrp_drugs <- unique(ctrp_ptmuts$Drug)
gdsc_drugs <- unique(gdsc_ptmuts$Drug)

@@

```{r histology_dists_data, eval=FALSE}
ccle_hists <- unique(select(ccle_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ccle_hists$Perc <- ccle_hists$n / sum(ccle_hists$n) * 100
ccle_hists$Dataset <- "CCLE"
ctrp_hists <- unique(select(ctrp_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
ctrp_hists$Perc <- ctrp_hists$n / sum(ctrp_hists$n) * 100
ctrp_hists$Dataset <- "CTRP"
gdsc_hists <- unique(select(gdsc_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
gdsc_hists$Perc <- gdsc_hists$n / sum(gdsc_hists$n) * 100
gdsc_hists$Dataset <- "GDSC"
crispr_hists <- unique(select(crispr_ptmuts, Broad_ID, Primary.Disease)) %>% count(Primary.Disease)
crispr_hists$Perc <- crispr_hists$n / sum(crispr_hists$n) * 100
crispr_hists$Dataset <- "CRISPR"

@@

```{r, eval=FALSE}
ccle_ccls <- unique(ccle_ptmuts$Broad_ID)
ctrp_ccls <- unique(ctrp_ptmuts$Broad_ID)
gdsc_ccls <- unique(gdsc_ptmuts$Broad_ID)
crispr_ccls <- unique(crispr_ptmuts$Broad_ID)

@@

```{r wilcoxon_tests, eval=FALSE}
crispr_ptmuts <- unfactorize(crispr_ptmuts)
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

@@

```{r mut_heatmaps, eval=FALSE}
ht_ccle <- MakeMutHeatmap(signif_g2p_druggene = ccle_signif_g2p_druggene, legend_title = "P-value", row_names = TRUE, color = "darkgreen", col_title = "Drugs")
png(filename = "./plots/manuscript/ccle_mut_heatmap.png", width = 3, height = 10, units = "in", res = 500)
draw(ht_ccle, heatmap_legend_side = "top")
dev.off()

@@

```{r drug_spear_calc, warning=FALSE, eval=FALSE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

@@

```{r drug_spear_plots, eval=FALSE}
# Copy number
ccle_cor <- unfactorize(filter(g2p_cor, Dataset == "CCLE"))
ccle_cn_ht <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.25, 0, 0.25), legend_title = "CCLE Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/ccle_cn_heatmap.png", width = 3, height = 10, units = "in", res = 500)
draw(ccle_cn_ht, heatmap_legend_side = "top")
dev.off()

@@

ccle_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
ctrp_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = ctrp_cor, row_names = FALSE, col_title = "CTRP", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
gdsc_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = gdsc_cor, row_names = FALSE, col_title = "GDSC", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
png(filename = "./plots/manuscript/drug_ge_heatmap.png", width = 13, height = 10, units = "in", res = 500)
draw(ccle_ge_ht2 + ctrp_ge_ht2 + gdsc_ge_ht2, gap = unit(0.25, "in"), heatmap_legend_side = "top")
dev.off()
```

@@

Kolmogorov-Smirnov tests comparing DGLs that are in a G2P DGL and not.
```{r drug_KS_tests, eval=FALSE, warning=FALSE, fig.height=6, fig.width=12}
# g2p_dgl_corr <- G2PCorrTable(ccle_ptmuts_df = ccle_ptmuts, ctrp_ptmuts_df = ctrp_ptmuts, gdsc_ptmuts_df = gdsc_ptmuts)
# saveRDS(g2p_dgl_corr, "./data_munging/rds/g2p_dgl_corr.rds")
g2p_dgl_corr <- readRDS("./data_munging/rds/g2p_dgl_corr.rds")
g2p_mut_stats <- G2PCorrStatsTables(corr_df = g2p_dgl_corr, table_type = "mutation")

g2p_corr_4plot <- filter(g2p_dgl_corr, G2P_Drug_Gene == "Yes")
g2p_corr_stats <- G2PCorrStatsTables(corr_df = g2p_corr_4plot, table_type = "KS")
g2p_corr_4plot$p <- NULL
g2p_corr_4plot$Mut_Signif <- NULL
g2p_corr_4plot <- melt(g2p_corr_4plot, id.vars = colnames(g2p_corr_4plot)[1:6], measure.vars = colnames(g2p_corr_4plot)[7:8], value.name = "Correlation")
g2p_corr_4plot <- g2p_corr_4plot %>% separate(col = variable, into = c("Correlation_Method", "corr", "Metric"), sep = "_", remove = FALSE)
g2p_corr_4plot$Metric <- ifelse(g2p_corr_4plot$Metric == "GE", "Gene Expression", "Copy Number")
